Geometric models reveal the hidden structure of conceptual knowledge

### 2023-01-19 paper/main.tex
added more references to the introduction

\author{Paxton C. Fitzpatrick\textsuperscript{1} and Jeremy R.
Manning\textsuperscript{*, 1}\\\textsuperscript{1}Dartmouth
College\\\textsuperscript{*}Corresponding author:
jeremy.r.manning@dartmouth.edu}

@@

How do we acquire conceptual knowledge? Memorizing course lectures or textbook
chapters by rote can lead to the superficial \textit{appearance} of
understanding the underlying content, but achieving true conceptual
understanding seems to require something deeper and richer. Does conceptual
understanding entail connecting newly acquired information to the scaffolding
of one's existing knowledge or experience? Or weaving a lecture's atomic
elements (e.g., its component words) into a structured network that describes
how those individual elements are related? Conceptual understanding could also
involve building a mental model that transcends the meanings of those
individual atomic elements by reflecting the deeper meaning underlying the
gestalt whole.

The difference between ``understanding'' and ``memorizing,'' as framed by the
researchers in education, cognitive psychology, and cognitive
neuroscience~\citep{Kato40, Gall10, ScotEtal13, HallGree08, Macl05} has
profound analogs in the fields of natural language processing and natural
language understanding. For example, considering the raw contents of a document
(e.g., its constituent symbols, letters, and words) might provide some
information about what the document is about, just as memorizing a passage
might be used to answer simple questions about the passage~\citep[e.g., whether
it might contain words related to furniture versus physics;][]{LandDuma97,
BleiEtal03, BleiLaff06}. However, modern natural language processing
models~\citep[e.g.,][]{MikoEtal13a, CerEtal18, BrowEtal20} also attempt to
capture the deeper meaning \textit{underlying} those atomic elements. These
models consider not only the co-occurrences of those elements within and across
documents, but also patterns in how those elements appear across different
scales (e.g., sentences, paragraphs, chapters, etc.), the temporal and
grammatical properties of the elements, and other high-level characteristics of
how they are used~\citep{Mann20, Mann21a}. According to these models, the deep
conceptual meaning of a document may be captured by a feature vector in a
high-dimensional representation space, where nearby vectors reflect
conceptually related documents. A model that succeeds at capturing an analog of
``understanding'' is able to assign nearby feature vectors to two conceptually
related documents, \textit{even when the words contained in those documents
have very little overlap}.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.8\textwidth]{figs/content-mastery}
\caption{\textbf{Caption.} Placeholder text.}
\label{fig:demo}
\end{figure}

>>>>

\author{Paxton C. Fitzpatrick and Jeremy R.
Manning\textsuperscript{*}\\Dartmouth
College\\\textsuperscript{*}Corresponding author:
jeremy.r.manning@dartmouth.edu}

@@

How do we acquire conceptual knowledge? Memorizing course lectures or textbook
chapters by rote can lead to the superficial \textit{appearance} of
understanding the underlying content, but achieving true conceptual
understanding seems to require something deeper and richer. Does conceptual
understanding entail connecting newly acquired information to the scaffolding
of one's existing knowledge or experience~\citep{BlayEtal06,CaraMaho03,
ConsEtal16, DeacEtal04, SimoEtal04}? Or weaving a lecture's atomic
elements (e.g., its component words) into a structured network that describes
how those individual elements are related? Conceptual understanding could also
involve building a mental model that transcends the meanings of those
individual atomic elements by reflecting the deeper meaning underlying the
gestalt whole~\citep{Kint70, Macl05, ScotEtal07}.

The difference between ``understanding'' and ``memorizing,'' as framed by the
researchers in education, cognitive psychology, and cognitive
neuroscience~\citep{Kato40, Gall00, ScotEtal07, HallGree08, Macl05} has
profound analogs in the fields of natural language processing and natural
language understanding. For example, considering the raw contents of a document
(e.g., its constituent symbols, letters, and words) might provide some
information about what the document is about, just as memorizing a passage
might be used to answer simple questions about the passage~\citep[e.g., whether
it might contain words related to furniture versus physics;][]{LandDuma97,
BleiEtal03, BleiLaff06}. However, modern natural language processing
models~\citep[e.g.,][]{MikoEtal13a, CerEtal18, BrowEtal20} also attempt to
capture the deeper meaning \textit{underlying} those atomic elements. These
models consider not only the co-occurrences of those elements within and across
documents, but also patterns in how those elements appear across different
scales (e.g., sentences, paragraphs, chapters, etc.), the temporal and
grammatical properties of the elements, and other high-level characteristics of
how they are used~\citep{Mann20, Mann21a}. According to these models, the deep
conceptual meaning of a document may be captured by a feature vector in a
high-dimensional representation space, where nearby vectors reflect
conceptually related documents. A model that succeeds at capturing an analog of
``understanding'' is able to assign nearby feature vectors to two conceptually
related documents, \textit{even when the words contained in those documents
have very little overlap}.

@@

% \begin{figure}[tp]
% \centering
% \includegraphics[width=0.8\textwidth]{figs/content-mastery}
% \caption{\textbf{Caption.} Placeholder text.}
% \label{fig:demo}
% \end{figure}
